\section{Limitations and Future Work}
\label{sec:design-and-future-work}
CodeInk's DM language supports traces of sorting and search algorithms across
lists, binary trees, and graphs. These algorithms comprise roughly one third of
the algorithms in CLRS~\cite{Cormen2001}: comparison sorts, operations on binary
search and red-black trees, breadth-first and depth-first graph traversal and
some shortest path algorithms. To provide completeness for the educational
domain, we plan to extend CodeInk's vocabulary to support additional data
structures (strings, linked lists, hash tables, and 2D arrays), as well as
compound data structures (lists of nodes, for example). This would enable traces
of a broader class of algorithms (dynamic programming, matrix operations), and
more complex traces (maintaining a queue of nodes in a breadth-first traversal).

Importantly, there is a distinction between tracing algorithms and describing
general programs. CodeInk's current focus is on the former, which is well-suited
for teaching algorithms via concrete traces. However, it is essential for
students to build an understanding of control abstraction: iteration sequences,
branching conditions, and recursion. We plan to extend CodeInk's vocabulary to
include gestures for specifying flow control. For example, we are prototyping
gestures for describing iteration by tapping objects and controlling the loop's
execution by dragging a slider between those objects. Our vision is to bring
CodeInk beyond descriptions of algorithm traces, to bridge the gap from concrete
examples to abstract program behavior.

